---
site: bookdown::bookdown_site
title: 'Exploration of Multi-Response Multivariate Methods'
alt-title: 'Utforskning av multi-respons multivariate metoder'
date: "`r format(Sys.Date(), format = '%m, %Y')`"
year: "`r format(Sys.Date(), format = '%Y')`"
month: "`r format(Sys.Date(), format = '%b')`"
location: 'Ås'
docsite: 'https://therimalaya.github.com/thesis'
author:
  name: "Raju Rimal"
  email: 'raju.rimal@nmbu.no'
  homepage: 'https://www.mathatistics.com/'
  affiliation: |
    Biostatistics\
    Dept. of Chemistry, Biotechnology and Food Science\
    Norwegian University of Life Sciences
supervisors:
  - title: 'Professor'
    name: 'Solve Sæbø'
    affiliation: |
      Prorector of Education\
      Norwegian University of Life Sciences\
      Ås, Norway
  - title: 'Associate Professor'
    name: 'Trygve Almøy'
    affiliation: |
      Dept. of Chemistry, Biotechnology and Food Science\
      Norwegian University of Life Sciences\
      Ås, Norway
bibliography: [References.bib]
biblio-style: 'plainnat'
biblio-title: 'References'
natbiboptions: 'authoryear'
nociteall: true
nocite: '@*'
#toc: true
lof: true
lot: false
knit: 'bookdown::render_book'
logo: "Logo.pdf"
logo-width: '0.6\linewidth'
thesis-number: "2019:76"
issn: "1894-6402"
isbn: "978-82-575-1636-9"
quote: "The goal is to turn data into information, and information into insight."
quote-author: "- Carly Fiorina, former CEO of Hewlett-Packard"
keywords: ['model-comparison ', 'multi-response ', 'simrel ']
acknowledgments: chapters/tex/00-Acknowledgment.tex
summary: chapters/tex/00-Summary.tex
sammendrag: chapters/tex/00-Sammendrag.tex
preface: chapters/tex/00-Preface.tex
papers:
  - title: "A tool for simulating multi response linear model data"
    path: "papers/001.pdf"
  - title: "Model and estimators for partial least squares regression"
    path: "papers/002.pdf"
  - title: "Comparison of Multi-response Prediction Methods"
    path: "papers/003.pdf"
  - title: "Comparison of Multi-response Estimation Methods"
    path: "papers/004.pdf"
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  comment    = NA,
  out.width  = '100%',
  dev        = ifelse(knitr::is_html_output(), "svg", "pdf"),
  fig.retina = 2,
  fig.pos    = "!htb",
  echo       = FALSE
)
```

```{r, include=FALSE}
source("_scripts/00-function.r")
load("../04-estimation-comparison/scripts/robj/design.rdata")
```

```{r load-pkg, include=FALSE}
library(tidyverse)
library(simrel)
library(pls)
```

# Introduction #

Rapid development in technology and computational power have resulted in heaps of data. Extracting information from this chaotic heaps of data has become another problem. Many statistical and machine learning tools are devised for this purpose, most of which focus to identify the relationships between different variables. A linear relationship is the most common assumption. This thesis confined itself in the exploration of linear relationships, where a set of independent variables, called predictor variables, affect another set of dependent variables, called response variables. The space spanned by the columns of predictors and responses are termed _predictor space_ and _response space_, respectively. 

Many projection-based statistical methods such as Principal Components Regression (PCR), Partial Least Squares (PLS) Regression and some variants of Envelopes only consider a subspace of predictor space relevant for defining the linear relationship between the predictors and the response(s). This brings us to the concept of relevant and irrelevant space introduced by @Naes1985. The relevant space can be described as the subspace that contains all the required information to define the relationship between the predictors and the response in a model. The irrelevant space, on the other hand, does not contain any information regarding this relationship. 

Latent components corresponding to predictor variables, which we will refer to as "predictor components", are linear combinations of the predictor variables. @Naes1985 and later @helland1990partial, @naes1993relevant and @Helland1994b have defined a set of predictor components as irrelevant components if they have no correlation with the response variables and the relevant part. Using only a subset of the latent components for modeling, is often termed as "dimension reduction". Methods like PCR, PLS and many other variants of PLS has leveraged this concept and are serving as prime tools in many disciplines, most notably in chemometrics.

Relatively new methods based on the concept of "envelopes" introduced by @Cook2007a, more specifically envelope in predictor variable (Xenv), have also used this concept of dimension reduction. In addition, envelope in response variable (Yenv) and simultaneous envelope in predictor and response (Senv) have extended the concept of relevant and irrelevant space to the response space as well, which they referred to as material and immaterial part. These methods are discussed in [Background] section.

Despite having similar underlying population model, these methods estimate the model parameters differently. Model parameters are the unknowns, which help to define a complex relationships between the variables. Regression coefficient vector $(\beta)$ in \@ref(eq:linear-reg-model) is an example of a model parameter. All methods use data to estimate these parameters. So, the properties of a dataset affect the estimation and consequently the prediction performance of the methods. Evaluation of these methods is essential to understand how they interact with various properties of the data. This thesis will explore some of these methods and assess their estimative and predictive strength and weaknesses through both simulated and real datasets.

This exploration adds a reference for researchers to motivate them for using different methods based on the properties of the data they are working on. This study is exploratory in nature where we assess and compare different multi-response multivariate methods, but most importantly study their interaction with the properties of the data. The properties include the correlation between predictor variables, the position of principal components of predictor variables (predictor components) that are relevant for certain principal components of the response variables (response components), the amount of correlation between the response variables and the number of predictor variables. The effect of the correlation structure of the response matrix is less explored and it is expected to shed some light on how similar and how different the methods are in terms of modelling this structure. In order to simulate data with these properties varying at different levels, we have created an R-package called `simrel`, which is an extension of the previous version introduced by @saebo2015simrel to incorporate multiple responses.

