---
site: bookdown::bookdown_site
title: 'Exploration of Multi-Response Multivariate Methods'
alt-title: 'Utforskning av multi-respons multivariate metoder'
date: "`r format(Sys.Date(), format = '%m, %Y')`"
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  comment    = NA,
  out.width  = '100%',
  dev        = c('pdf', 'svg'),
  fig.retina = 2,
  fig.pos    = "!htb",
  echo       = FALSE
)
```


# Introduction

Try to be precise on the thesis objective, problem and what achievement does the theses is expected to meet.

- What is the thesis about
- Why is this important
- What will it add to the big picture


```{r a-test-figure, fig.cap="Test Figure", fig.width=5, fig.asp=0.7}
library(ggplot2)
ggplot(mtcars, aes(hp, mpg, color = factor(am))) +
    geom_point(aes(size = cyl))
```

```{r a-test-table}
knitr::kable(
  x        = head(mtcars[, 1:7]),
  caption  = "The head of mtcars",
  booktabs = TRUE
)
```

## Background

Write very basic about following topics as an introductory text.

- About reduction of regression model
- Relevant and irrelevant components
- Various study up to this point
- Multi-response model
- Tool for simulation


---
year: "`r format(Sys.Date(), format = '%Y')`"
month: "`r format(Sys.Date(), format = '%b')`"
location: 'Ås'
docsite: 'https://therimalaya.github.com/thesis'
author:
  name: "Raju Rimal"
  email: 'raju.rimal@nmbu.no'
  homepage: 'https://www.mathatistics.com/'
  affiliation: |
    Biostatistics\
    Dept. of Chemistry, Biotechnology and Food Science\
    Norwegian University of Life Sciences
supervisors:
  - title: 'Professor'
    name: 'Solve Sæbø'
    affiliation: |
      Prorector\
      Norwegian University of Life Sciences\
      Ås, Norway
  - title: 'Associate Professor'
    name: 'Trygve Almøy'
    affiliation: |
      Dept. of Chemistry, Biotechnology and Food Science\
      Norwegian University of Life Sciences\
      Ås, Norway
bibliography: 'References.bib'
biblio-style: 'plainnat'
biblio-title: 'References'
natbiboptions: 'authoryear'
nociteall: true
toc: true
lof: true
lot: true
knit: 'bookdown::render_book'
logo: "Logo.pdf"
logo-width: '0.6\linewidth'
thesis-number: "1234:56"
issn: "1234-5678"
isbn: "123-45-678-1234-5"
quote: "The goal is to turn data into information, and information into insight."
quote-author: "- Carly Fiorina, former CEO of Hewlett-Packard"
keywords: ['model-comparison ', 'multi-response ', 'simrel ']
title-abstract: |
  This thesis is prepared with `ArsClassica` \LaTeX\ template with `pandoc` and r-package `bookdown`.\
  Source: [https://github.com/therimalaya/thesis](https://github.com/therimalaya/thesis)
acknowledgments: chapters/00-Acknowledgment.md
summary: chapters/00-Summary.md
sammendrag: chapters/00-Sammendrag.md
preface: chapters/00-Preface.md
# papers:
#   - title: "A tool for simulating multi-response linear model data"
#     path: "papers/001.pdf"
#   - title: "Model and estimators for partial least squares regression"
#     path: "papers/002.pdf"
#   - title: "Comparison of Multi-response Prediction Methods"
#     path: "papers/003.pdf"
---
